---
title: "Final Project"
author: "Shelby, Imani, Maitreyi, Hannah"
date: "5/1/2022"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Executive Summary
Our project was motivated by the need to understand the demand for and the supply of social services in the United States of America. To narrow this down further, we restricted this study to the state of California, as there is relatively greater access to social services such as WIC and SNAP in this state. In FY 2021, SNAP helped 4.3 million CA residents or 11% of the population, and in FY 2021, WIC helped 950,000 CA residents.

Studies also indicate that many households in California struggle to put food on the table. This statistic indicates the need for further analysis of WIC and SNAP services in California.

Most recent data indicates that:

* 9.8% of households were “food insecure"
* 11% of the population lived below the poverty line
* 14.5% of children lived below the poverty line
* 10.7% of older adults lived below the poverty line

Policy Questions:

* What does social service delivery look like in CA and how well is the state meeting demand?
* Where are services more accessible or less accessible to those who are eligible?
* Why are redemption rates of social services lower than 60% in CA?

We used data from California Open Data portals, [USDA.gov](https://www.usda.gov/), and census data as well to conduct our analyses.

We used a combination of tools to answer our motivating policy questions. First, we began by conducting exploratory analyses, including the use of graphs and geospatial visualizations, to further analyze access to WIC and SNAP store locations in different counties in California. We were able to discern trends between different populations and identify the impact of development (urban or rural county) on access to social services. We then used supervised and unsupervised machine learning models (text analysis, cluster analysis, linear regression model) to predict store locations that were likely to have WIC or SNAP and number of vouchers redeemed across counties, based on date and number of families in the region.

The challenges we came across while coding this project were largely related to the data we had access to. We were unable to conduct analysis by race and gender, because of the way these factors were coded into the table. With more time and access to resources, we would love to analyze how women and non-binary folks access these social services. Comparing and contrasting these results in urban counties and rural counties will also be critical to policy making.

## Data Loading and Cleaning
```{r, warning=FALSE, message=FALSE}
#loading and cleaning data in advance so that it can be used in a combination of plot below

library(tidycensus)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(janitor)
library(sf)
library(stringr)
library(lubridate)
library(ggridges)
library(viridis)
library(hrbrthemes)

#1. SNAP Store Locations - "retailers"
retailers <- read_csv("SNAP_Store_Locations.csv")
retailers <- retailers %>%
  janitor::clean_names() %>%
  filter(state == "CA") %>%
  mutate(COUNTY = toupper(county)) %>%
  select(-county, -address_line_2)

# creating urban and rural counties
urban <- c("LOS ANGELES",  "ORANGE", "SAN FRANCISCO", "SAN MATEO", "ALAMEDA", "CONTRA COSTA", "MARIN", "RIVERSIDE", 
"SACRAMENTO", "SAN BERNARDINO", "SAN DIEGO", "SAN JOAQUIN", "SANTA CLARA", "VENTURA", "FRESNO")

  retailers <- retailers %>%
  mutate(urban= if_else(COUNTY %in% urban, 1, 0))

  rural <-c("ALPINE", "AMADOR", "BUTTE", "CALAVERAS", "COLUSA", "DEL NORTE", "EL DORADO", "GLENN",
  "HUMBOLDT", "IMPERIAL", "INYO", "LAKE", "LASSEN", "MADERA", "MARIPOSA", "MENDOCINO", "MERCED", "MODOC",
  "MONO", "MONTEREY", "NAPA", "NEVADA", "PLACER", "PLUMAS", "SAN BENITO", "SAN LUIS OBISPO", 
  "SANTA BARBARA", "SHASTA", "SIERRA", "SISKIYOU", "SOLANO", "SONOMA", "SUTTER", "TEHAMA", "TRINITY",
  "TULARE", "TUOLUMNE", "YOLO", "YUBA", "STANISLAUS", "KERN", "KINGS", "SANTA CRUZ")

  retailers <- retailers %>%
  mutate(rural= if_else(COUNTY %in% rural, 1, 0))

#prepping for visualizations & joining to census data
#creating ca_retailers
ca_retails <- retailers %>%
  group_by(COUNTY) %>%
  summarize(N_RETAIL = n()) %>%
  ungroup()

#2. Census data + SNAP retailer data
county_pop_url <- "https://www2.census.gov/programs-surveys/popest/datasets/2010-2019/counties/totals/co-est2019-alldata.csv"
county_pop_raw <- read_csv(url(county_pop_url))
ca_pop <- county_pop_raw %>%
  filter(SUMLEV == "050") %>%
  filter(STNAME == "California") %>%
  select(
    COUNTY = CTYNAME,
    POPESTIMATE2019) %>%
  mutate(COUNTY = toupper(COUNTY),
         COUNTY = gsub("(.*)( COUNTY)", "\\1", COUNTY))

#combined data with retailers and pop data
combined_ca_data <- left_join(ca_retails, ca_pop, by = "COUNTY") %>%
  mutate(retailers_per_100k = N_RETAIL / (POPESTIMATE2019/100000))

#save retailers_per_100k as an object
retailers_per_100k <- combined_ca_data %>%
  mutate(retailers_per_100k = N_RETAIL / (POPESTIMATE2019/100000))

#3. SNAP participant demographic counties
snap_demos <- read_csv("particp_demog_counties.csv")
tidy_demos <- snap_demos %>%
  clean_names() %>%
  rename(year = fileyear) %>%
  mutate(COUNTY = toupper(county)) %>%
  select(year, COUNTY, person, cases, female, male, black, hispanic, asian_pi, native_american_other_unknown, white) %>%
  filter(COUNTY != "COUNTY TOTAL")
## issue with this dataset is that it does not include # of cases by those demographics. it's only showing demographics for the county and how many cases there are.
tidy_demos <- tidy_demos %>%
    mutate(rate = cases/person)

#4. Cal Fresh Data
calfresh <- read_csv("calfresh.csv")
tidy_calfresh <- calfresh %>%
  janitor::clean_names() %>%
  rename(COUNTY = county) %>%
  mutate(COUNTY = toupper(COUNTY))

#5. WIC vendors
wic_vendors <- read_csv("vendor.csv")
wic_tidy <- wic_vendors %>%
  mutate(COUNTY = toupper(COUNTY))


# WIC retailers data
wic_retailers <- wic_tidy %>%
  group_by(COUNTY) %>%
  summarize(N_RETAIL = n()) %>%
  ungroup()


#6 WIC redemption 

#by county
wic_redemp_county <- read_csv("wic_county.csv") %>%
  clean_names()

#by particpant
wic_redemp_particp <- read_csv("wic_participant.csv") %>%
  clean_names() %>%
  separate(year_month, c("year", "month")) %>%
  rename(COUNTY = vendor_location) %>%
  filter(COUNTY != "STATEWIDE ANNUAL") %>%
  select(-statewide_infant_formula_rebate, -total_cost_vouchers, -total_cost_vouchers_adjusted, -average_cost_adjusted,    
         -state_average_cost_adjusted) 

wic_redemp_particp <- wic_redemp_particp %>%
  mutate(across(starts_with("average"), ~gsub("\\$", "", .) %>% as.numeric)) %>%
  mutate(month = as.numeric(month)) %>%
  mutate(year = as.numeric(year))

wic_part <- wic_redemp_particp %>%
  group_by(year, COUNTY, participant_category)
wic_part$number_of_participants_redeemed <- as.numeric(gsub("," , "", wic_part$number_of_participants_redeemed))
wic_part$number_vouchers_redeemed <- as.numeric(gsub("," , "", wic_part$number_vouchers_redeemed))

wic_binary <- wic_redemp_particp %>% 
  mutate(urban = if_else(COUNTY %in% urban, 1, 0))

```

# Exploratory Data Analysis 
```{r, WIC, Average Cost Per Participant}

wic_redemp_particp%>%
  ggplot() +
  geom_col(mapping = aes(participant_category, average_cost, fill = participant_category)) +
  geom_text(
    mapping = aes(participant_category, average_cost, label = average_cost, angle=90, hjust= -0.3),
    stat = "identity",
    parse = FALSE,
    nudge_x = 0,
    nudge_y = 0,
    check_overlap = TRUE,
    na.rm = FALSE,
    show.legend = NA,
    inherit.aes = TRUE
    ) + 
  facet_wrap(~year) +
  labs(
    title = "Average Cost by WIC Participant",
    x = "Type of Participant",
    y = "Average Cost per Participant (in $)"
  ) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

``` 

```{r, WIC Participation Rates by Type}
wic_time_series <- read_csv("participant_rates.csv") %>%
  pivot_longer(cols=c("2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018"),
               names_to = "year",
               values_to = "rate")


wic_time_series %>%
  mutate(year = as.numeric(year)) %>%
  ggplot(mapping = aes(x = year, y = rate, color = participant_category)) +
  geom_line(alpha = 0.6, size = 1) +
  labs(
    title = "WIC Participation Rates by Type",
    x = "Year",
    y = "Rate")


```

```{r, EDA WIC Redemption}

#THIS IS BEAUTIFUL PLEASE LEAVE IT ALONE - See if we can adjust?
wic_redemp_particp %>%
  ggplot(mapping = aes(x=participant_category, y=average_cost, fill = participant_category)) +
  geom_violin() +
  xlab("class") +
  theme(legend.position = "none") +
  xlab("") +
  theme_minimal()

#flat and pointy - why?
wic_binary %>%
  ggplot(mapping = aes(x=participant_category, y=average_cost, fill = participant_category)) +
  geom_violin() +
  facet_wrap(~urban) +
  xlab("class") +
  theme(legend.position = "none") +
  xlab("") +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  labs(
    title = "Average Cost by Participant, 0=Rural and 1=Urban",
    source = "California Open Data"
  )

```

# Geospatial Analyses 
```{r, geospatial, warning=FALSE, message=FALSE, results=FALSE}

#Number of Retailers per 100,000 People by County in 2019
library(tigris)

#creating ca_counties
ca_counties_raw <- tigris::counties(
  state = "CA",
  cb = TRUE,
  resolution = "500k",
  year = 2020,
  class = "sf")

#getting geo outline of CA by county
ca_counties_raw %>%
  ggplot() +
  geom_sf() +
  theme_void()

ca_counties <- ca_counties_raw %>%
  dplyr::transmute(
    GEOID,
    MAP_NAME = NAME,
    COUNTY = toupper(NAME)
  )

#Geo spatial graph - Retailers per 100k People
ca_geospatial_data <- geo_join(
  spatial_data = ca_counties,
  data_frame = combined_ca_data,
  by_sp = "COUNTY",
  by_df = "COUNTY",
  how = "left")
ggplot(ca_geospatial_data, aes(fill = retailers_per_100k)) +
  geom_sf() +
  #scale_fill_viridis_c() +
  labs(
    title = "Number of Retailers per 100,000 People by County in 2019"
  ) +
  theme_void()

```

```{r, geospatial, Cal fresh: child, adult, and elderly participation rates}

#Cal fresh geospatial analysis for child, adult, and elderly participation rates
join_calfresh <- tidy_calfresh %>%
  select(COUNTY, year, elderly, adults, children, esl, total_population_cy, total_elderly_60plus_cy, total_children_under_18_cy, total_esl_over_age_5_cy) %>%
  na.omit()
join_calfresh <- join_calfresh %>%
  mutate(rate_elder = elderly/total_elderly_60plus_cy) %>%
  mutate(rate_child = children/total_children_under_18_cy) %>%
  mutate(rate_adult = adults/total_population_cy)
calfresh_map <- geo_join(
  spatial_data = ca_counties,
  data_frame = join_calfresh,
  by_sp = "COUNTY",
  by_df = "COUNTY",
  how = "inner")
map_child <- ggplot(calfresh_map, aes(fill = rate_child)) +
  geom_sf() +
  scale_fill_viridis_c() +
  theme_void()
map_adult <- ggplot(calfresh_map, aes(fill = rate_adult)) +
  geom_sf() +
  scale_fill_viridis_c() +
  theme_void()
map_elder <- ggplot(calfresh_map, aes(fill = rate_elder)) +
  geom_sf() +
  scale_fill_viridis_c() +
  theme_void()
library(patchwork)
map_child + map_adult + map_elder
```

```{r, geospatial, Cal fresh: child, adult, and elderly participation rates, by year}
#cal fresh map by year
calfresh_map %>%
  ggplot(mapping = aes(fill = rate_child)) +
  geom_sf() +
  scale_fill_viridis_c() +
  facet_wrap(~year) +
  labs(title = "Child SNAP Participation Rate") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
calfresh_map %>%
  ggplot(mapping = aes(fill = rate_elder)) +
  geom_sf() +
  scale_fill_viridis_c() +
  facet_wrap(~year) +
  labs(title = "Elderly SNAP Participation Rate") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
calfresh_map %>%
  ggplot(mapping = aes(fill = rate_adult)) +
  geom_sf() +
  scale_fill_viridis_c() +
  facet_wrap(~year) +
  labs(title = "Adult SNAP Participation Rate") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
```
This overall rate graph shows that there is a large percentage of eligible folks across California who do not participate in SNAP despite being eligible for services. The ideal rate, of course, is 100% providing service to all who are eligible. This begs the question: what can California do better to increase the participation rate? Who is not being reached?

Future research could include data that provides participation rate of women and of women who are parents to one or more children. 

```{r, geospatial, overall SNAP participation rate}
# SNAP Map
snap_map <- geo_join(
  spatial_data = ca_counties,
  data_frame = tidy_demos,
  by_sp = "COUNTY",
  by_df = "COUNTY",
  how = "inner")
snap_map %>%
  ggplot(mapping = aes(fill = rate)) +
  geom_sf() +
  scale_fill_viridis_c() +
  facet_wrap(~year) +
  labs(title = "SNAP Participation Rate") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

```


```{r, geospatial, WIC retailers per 100K}

#WIC retailers per 100K
wic_join <- left_join(wic_retailers, ca_pop, by = "COUNTY") %>%
  mutate(retailers_per_100k = N_RETAIL / (POPESTIMATE2019/100000))
library(tigris)
wic_geo_data <- geo_join(
  spatial_data = ca_counties,
  data_frame = wic_join,
  by_sp = "COUNTY",
  by_df = "COUNTY",
  how = "left")
ggplot(wic_geo_data, aes(fill = retailers_per_100k)) +
  geom_sf() +
  #scale_fill_viridis_c() +
  labs(
    title = "WIC Retailers per 100k"
  ) +
  theme_void()
```

## WIC Machine Learning
```{r, warning=FALSE}

library(tidymodels)
library(rsample)
library(parsnip)
library(recipes)
library(workflows)
library(tune)
library(yardstick)
library(glmnet)

#Data loading and cleaning
wic_models <- read_csv("wic_participant.csv") %>%
  janitor::clean_names() %>%
  filter(vendor_location != "STATEWIDE ANNUAL") %>%
  filter(vendor_location != "STATEWIDE") %>%
select(-statewide_infant_formula_rebate, -total_cost_vouchers_adjusted, -average_cost_adjusted, 
  -state_average_cost_adjusted, -total_cost_vouchers, -participant_category, -vendor_location) %>%
separate(year_month, c("year", "month")) %>%
  mutate(year = as.numeric(year)) %>%
  mutate(month = as.numeric(month))

wic_models$number_of_participants_redeemed <- as.numeric(gsub("," , "", 
  wic_models$number_of_participants_redeemed))
wic_models$number_vouchers_redeemed <- as.numeric(gsub("," , "", 
  wic_models$number_vouchers_redeemed))
wic_models$average_cost = as.numeric(gsub("\\$", "", wic_models$average_cost))


#predict number of vouchers redeemed
#----------------------------------------
#linear regression

#initial split
set.seed(20211102)
split <- initial_split(wic_models, prop = 0.7, strata = "number_vouchers_redeemed")
wic_train <- training(split)
wic_test <- testing(split)

#recipe
wic_rec <- recipe(number_vouchers_redeemed ~., data = wic_train) %>%
  step_nzv(all_predictors()) %>%
  step_normalize(all_predictors())

# set up resampling using 10-fold cross validation
set.seed(20211102)
folds <- vfold_cv(data = wic_train, v = 10, repeats = 1)

# create a linear regression model
wic_mod <- linear_reg() %>%
  set_engine("lm")

# create a workflow 
wic_wf <- workflow() %>%
  add_recipe(wic_rec) %>%
  add_model(wic_mod)

# fit the model 
wic_cv <- wic_wf %>%
  fit_resamples(resamples = folds)

# select the best model
wic_best <- wic_cv %>%
  select_best("rmse")

# finalize the workflow
wic_final <- finalize_workflow(
  wic_wf,
  parameters = wic_best
)

# fit to the training data and extract coefficients
wic_coefs <- wic_final %>%
  fit(data = wic_train) %>%
  extract_fit_parsnip() %>%
  vip::vi(lambda = wic_best$penalty)

collect_metrics(wic_cv, summarize=TRUE)%>%
  filter(.config == "Preprocessor1_Model1")

#plot the RMSE across the 10 resamples
collect_metrics(wic_cv, summarize = FALSE) %>%
filter(.config == "Preprocessor1_Model1",
       .metric == "rmse") %>%
ggplot(aes(id, .estimate, group = .metric)) +
geom_line() +
geom_point() +
scale_y_continuous() +
labs(title = "Calculated RMSE Across the 10 Folds",
y = "RMSE_hat") +
theme_minimal()

#fit to test data
wic_final_model <- fit(wic_wf, data = wic_test)

#make predictions
final_predictions <- bind_cols(
  wic_models,
  predict(wic_final_model, wic_models,
          type="numeric"))

print(final_predictions)

#----------
#Lasso regression

#initial split
#set.seed(20211101)
#lasso_split <- initial_split(data = wic_models, prop = 0.75)
# create the training and testing data
#lasso_train <- training(x = lasso_split)
#lasso_test <- testing(x = lasso_split)

#folds
#l_folds <- vfold_cv(data = lasso_train, v = 10)

#Create wic recipe
#wic_rec <- recipe(average_cost ~., data = wic_train) %>%
#  step_nzv(all_predictors()) %>%
 # step_normalize(all_predictors()) %>%
#  step_indicate_na()

#tuning grid for lasso
#lasso_grid <- grid_regular(penalty(), levels = 10)

#lasso mod
#lasso_mod <- 
 # linear_reg(
 #   penalty = tune(),
 #   mixture = 1
 # ) %>%
# 3 set_engine(engine = "glmnet") 

#lasso workflow
#lasso_wf <- 
 #   workflow() %>%
 #   add_model(spec = lasso_mod) %>%
 #   add_recipe(recipe = wic_rec)

#hyperparameter tuning
#lasso_cv <- lasso_wf %>%
 #   tune_grid(
  #    resamples = folds,
  #    grid = lasso_grid)

# select the best model based on the "rmse" metric
#lasso_best <- lasso_cv %>%
 # select_best(metric = "rmse")

# use the finalize_workflow() function with your lasso workflow and the best model
# to update (or "finalize") your workflow by modifying the line below
#lasso_final <- finalize_workflow(
#  lasso_wf,
 # parameters = lasso_best)

# fit to the training data and extract coefficients
#lasso_coefs <- lasso_final %>%
#  fit(data = chi_train) %>%
#  extract_fit_parsnip() %>%
 # vip::vi(lambda = lasso_best$penalty)

#collect_metrics(lasso_cv, summarize=TRUE) %>%
 # filter(.config == "Preprocessor1_Model10")

#plot the RMSE across the 10 resamples
#collect_metrics(lasso_cv, summarize = FALSE) %>%
#filter(.config == "Preprocessor1_Model10",
     #  .metric == "rmse") %>%
#ggplot(aes(id, .estimate, group = .metric)) +
#geom_line() +
#geom_point() +
#scale_y_continuous() +
#labs(title = "Calculated RMSE Across the 10 Folds",
#y = "RMSE_hat") +
#theme_minimal()



#  mutate(across(starts_with("average"), ~gsub("\\$", "", .) %>% as.numeric)) %>%
#  mutate(across(starts_with("total"), ~gsub("\\$", "", .) %>% as.numeric))
 # select(participant_category, number_of_participants_redeemed, number_vouchers_redeemed)
#wic_models$total_cost_vouchers = as.numeric(gsub("\\$", "", wic_models$total_cost_vouchers))


```

# Cluster Analysis

** Add to interpretation: Squares that are perfectly white tell us that there is no correlation between cost and year, indicating that the cost is consistent across years. Number of participants and number of vouchers are perfectly correlated, indicating that there is not a difference across people in the number of vouchers that participants receive. 

```{r, Convert WIC data to numeric datafram, create correlation matrix}
library(corrplot)

#create numeric data for PCA analysis
wic_numeric <- wic_part %>%
  filter(COUNTY == "STATEWIDE") %>%
  select(number_of_participants_redeemed, number_vouchers_redeemed, average_cost) %>%
  mutate(part_type = case_when(
    participant_category == "Breastfeeding Mother" ~ "1",
    participant_category == "Child" ~ "2",
    participant_category == "Infant" ~ "3",
    participant_category == "Non-Breastfeeding Mother" ~ "4",
    participant_category == "Prenatal" ~ "5"
  )) %>%
  ungroup() %>%
  select(-participant_category, -COUNTY) %>%
  mutate(part_type = as.numeric(part_type)) %>%
  rename(n_vouchers = number_vouchers_redeemed) %>%
  rename(n_part = number_of_participants_redeemed)

correlation_plot <- cor(wic_numeric)
# creates a correlation plot
corrplot(correlation_plot, method = "shade")

```

```{r, PCA Extraction and Column Binding, warning=FALSE}

wic_state <- wic_part %>%
   filter(COUNTY == "STATEWIDE") %>%
  select(number_of_participants_redeemed, number_vouchers_redeemed, average_cost) %>%
  mutate(part_type = case_when(
    participant_category == "Breastfeeding Mother" ~ "1",
    participant_category == "Child" ~ "2",
    participant_category == "Infant" ~ "3",
    participant_category == "Non-Breastfeeding Mother" ~ "4",
    participant_category == "Prenatal" ~ "5"
  )) %>%
  ungroup() %>%
  mutate(part_type = as.numeric(part_type)) %>%
  rename(n_vouchers = number_vouchers_redeemed) %>%
  rename(n_part = number_of_participants_redeemed)
wic_pca_numeric <- wic_state %>%
  select_if(is.numeric)
#run PCA
pca_wic <- prcomp(wic_pca_numeric)
#extract principle components
wic_pcs <- pca_wic %>%
  .$x %>%
  as_tibble()
#combine pcs to county and participant category
wic_pcs <- bind_cols(
  select(wic_state, COUNTY, participant_category),
  wic_pcs
)
```

```{r, WIC cluster analysis, warning=FALSE} 

library(cluster)
library(dendextend)
library(factoextra)
library(GGally)
library(recipes)
# create a recipe with no outcome variable and all predictors
wic_pca_rec <- recipe(~., data = wic_numeric) %>%
  # center and scale all predictors
  step_center(all_predictors()) %>%
  step_scale(all_predictors()) %>%
  # run prep to prepare recipe
  prep()
# apply recipe to employment_rec
wic_clust <- wic_pca_rec %>%
  bake(new_data = NULL)
# PCA ---------------------------------------------------------------------
# create a correlation matrix on employment_clust
cor(wic_clust)
# conduct PCA on the employment_clust data
principal_components <- prcomp(wic_clust)
# obtain summary metrics
summary(principal_components)
# obtain loadings
principal_components$rotation
# obtain component values for each observation
pca_data <- as_tibble(principal_components$x) %>%
  select(PC1, PC2, PC3)

# set a seed because the clusters are not deterministic
set.seed(20200205)
# total within sum of squares
fviz_nbclust(wic_clust, FUN = kmeans, method = "wss")
# total silhouette width
fviz_nbclust(wic_clust, FUN = kmeans, method = "silhouette")
# gap statistic
fviz_nbclust(wic_clust, FUN = kmeans, method = "gap_stat")
# run kmeans with the optimal number of clusters using the employment_clust data, set nstart = 100
clust_kmeans <- kmeans(
  wic_clust,
  centers = 6,
  nstart = 100
)
# examine the cluster means
tidy(clust_kmeans)

```

```{r continuing class tutorial}

#fit with 6 clusters
wic_kmeans6 <- kmeans(
  wic_pca_numeric,
  centers = 6,
  nstart = 100
)
bind_cols(
  select(wic_state, COUNTY, participant_category),
  cluster = wic_kmeans6$cluster
) %>%
  count(participant_category, cluster)

wic_clusters <- bind_cols(
  select(wic_state, COUNTY, participant_category),
  select(wic_pcs, PC1, PC2, PC3, PC4, PC5),
  cluster6 = wic_kmeans6$cluster)
ggplot() +
  geom_point(
    data = wic_clusters,
    mapping = aes(PC1, PC2, color = factor(cluster6)),
    alpha = 0.5) +
  theme_minimal()

```


## Text Analysis - SNAP Store Locations

```{r, Text Analysis, warning=FALSE, message=FALSE}

library(tidytext)
library(igraph)
library(ggraph)

#use retailers data that was cleaned earlier
retailers_text <- retailers %>% 
  filter(!is.na(store_name)) %>%
  select(-zip4)

#creating biagram
bigram_snap <- retailers %>%
  unnest_tokens(bigram, store_name, token = "ngrams", n = 2)

#separating the bigram into two columns
bigrams_separated <- bigram_snap %>%
  separate(bigram, c("word1", "word2"), sep = " ")

#filtering out rows without stopwords
bigrams_filtered <- bigrams_separated %>%
  filter(!word1 %in% stop_words$word & !word2 %in% stop_words$word)

#counting number of appearances of each bigram and filtering the rows
bigram_30 <- bigrams_filtered %>%
  count(word1, word2, sort = TRUE) %>%
  filter(n > 30) %>%
  filter(!is.na(word1)) 

# plot the bigrams that exist more than 30 times
bigram_graph <- bigram_30 %>%
  graph_from_data_frame()
# plot the relationships
set.seed(2017)
ggraph(bigram_graph, layout = "fr") +
  geom_edge_link() +
  geom_node_point() +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1)
# merging to a single column
combined1 <- bigrams_filtered %>%
  mutate(bigram = paste(word1,word2))

#-------------------------
#repeat geom_col for the 6 largest urban counties, 6 largest rural/suburban, and 6 smallest rural

#create stopword vectors for city names that don't tell us anything about the name/type of store
urb_city_names <- c("santa", "clara", "ana", "san", "jacinto", "monica", "los", "angeles", "pedro", "monte",
                    "palma", "chula", "vista", "cajon", "quinta", "jose", "yucca", "huntington", "diego", "ysidro",
                    "costa", "mesa", "lucerne", "grove")

rural_city_names <- c("santa", "fe", "central", "wheeler")


# filtering for the 6 largest urban counties
tf_idf_urb <- combined1 %>%
  filter(COUNTY == "LOS ANGELES" | COUNTY == "SAN DIEGO" | COUNTY == "ORANGE" | COUNTY == "RIVERSIDE" | COUNTY == "SAN BERNARDINO" | COUNTY == "SANTA CLARA") %>%
  filter(!word1 %in% stop_words$word,
         !word2 %in% stop_words$word,
         !word1 %in% urb_city_names,
         !word2 %in% urb_city_names,
         !is.na(word1), !is.na(word2)) %>%
  count(COUNTY, bigram) %>%
  bind_tf_idf(bigram, COUNTY, n) %>%
  #plotting the 10 most frequent
  group_by(COUNTY) %>%
  slice_max(order_by=tf_idf, n=10) %>%
  mutate(bigram=reorder(bigram, tf_idf)) %>%
  ggplot() +
  geom_col(aes(tf_idf, bigram, fill=COUNTY)) +
  facet_wrap(~COUNTY, scales="free")+
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  guides(fill="none") 

tf_idf_urb


# - - - - - - - - - -
#filter for 6 largest rural counties
tf_idf_rur_high <- combined1 %>%
  filter(COUNTY == "KERN" | COUNTY == "STANISLAUS" | COUNTY == "SONOMA" | COUNTY == "TULARE" | COUNTY == "SOLANO" | COUNTY == "MONTEREY") %>% #chose not to use Santa Barbara because it did not feel reflective of rural
   filter(!word1 %in% stop_words$word,
         !word2 %in% stop_words$word,
         !word1 %in% rural_city_names,
         !word2 %in% rural_city_names,
         !is.na(word1), !is.na(word2)) %>%
  count(COUNTY, bigram) %>%
  bind_tf_idf(bigram, COUNTY, n) %>%
  # plotting the 5 most frequent store names
  group_by(COUNTY) %>%
  slice_max(order_by=tf_idf, n=5) %>%
  mutate(bigram=reorder(bigram, tf_idf)) %>%
  ggplot(aes(tf_idf, bigram, fill=COUNTY)) +
  geom_col() +
  facet_wrap(~COUNTY, scales="free")+
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  guides(fill="none") +
  scale_x_continuous()

tf_idf_rur_high

# - - - - - - - - - -
#filter for 6 smallest rural counties
tf_idf_rur_low <- combined1 %>%
  filter(COUNTY == "ALPINE" | COUNTY == "SIERRA" | COUNTY == "MODOC" | COUNTY == "AMADOR" | COUNTY == "MONO" | COUNTY == "MARIPOSA") %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word) %>%
  count(COUNTY, bigram) %>%
  bind_tf_idf(bigram, COUNTY, n) %>%
  # plotting the 5 most frequent store names
  group_by(COUNTY) %>%
  slice_max(order_by=tf_idf, n=5) %>%
  mutate(bigram=reorder(bigram, tf_idf)) %>%
  ggplot(aes(tf_idf, bigram, fill=COUNTY)) +
  geom_col() +
  facet_wrap(~COUNTY, scales="free")+
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  guides(fill="none")

tf_idf_rur_low

```













